---
title: "Lab 04"
author: "Alex Connolly"
date: "Feb-16-2022"
output: github_document
---

### Load packages and data

```{r load-packages}
library(tidyverse) 
library(dsbox) 
```

```{r load-data}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r dennys}
nrow(dennys)
ncol(dennys)

```

There are 1643 rows and 6 columns, the variables are addresses, city, states, zips, longitude, latitude.
### Exercise 2

```{r laquinta}
nrow(laquinta)
ncol(laquinta)

```
909 rows, 6 columns, same variables

### Exercise 3

There are multiple La Quinta locations outside of the US, including Canada, Mexico, China, New Zealand, Honduras, Turkey, United Arab Emirates,Chile, Colombia.

No Dennys in other countries. 

### Exercise 4

You could filter locations that are from outside the US. 


### Exercise 5
```{r statesdennys}
dennys %>%
  filter(!(state %in% states$abbreviation))
```

No states are outside the US for Dennys

### Exercise 6

```{r countrydennys}
dennys %>%
  mutate(country = "United States")
```

### Exercise 7
 
```{r statelaquinta}
laquinta %>%
  filter(!(state %in% states$abbreviation))
```
These countries ar Mexico (10 of them), Colombia, Canada (2), and Hondurus. 

### Exercise 8

```{r countries}
laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state == "FM"            ~ "Honduras",
    state %in% c("AG", "QR", "CH", "NL", "VE", "PU", "SL") ~ "Mexico"
  ))
```
```{r filterUS}
laquinta %>%
  filter(country == "United States")
```

### Excercise 9



```{r joiningD}

dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(Dennys_per_distance = n / area * 1000) %>%
  arrange(Dennys_per_distance)

```
DC has the most Dennys per thousand square miles, rhode island second.
Delaware has the least Dennys, California has the most.

```{r joiningL}

laquinta %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation")) %>%
  mutate(Laquinta_per_distance = n / area * 1000) %>%
  arrange(Laquinta_per_distance)

```
Here, RI has the most per thousand square miles, FLorida second.
Maine has the least La Quintas, Texas has the most. 
 
### Excersise 10

```{r mutate-bind}
dennys <- dennys %>%
  mutate(establishment = "Denny's")
laquinta <- laquinta %>%
  mutate(establishment = "Laquinta")
dennys_laquinta <- bind_rows(dennys, laquinta)
```

```{r plot}
ggplot(dennys_laquinta, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```
### Excersise 11
```{r North_Carolian}
dennys_laquinta %>%
  filter(state == "NC") %>%
  ggplot(dennys_laquinta, mapping = aes(x = longitude, y = latitude, color = establishment)) + geom_point(alpha=.5) 
```
I mean there are definitely a lot of Denny's that are close to Laquintas, but I don't think it is as close as Mitch Hedberg suggests.

### Excercise 12

```{r Texas}
dennys_laquinta %>%
  filter(state == "TX") %>%
  ggplot(dennys_laquinta, mapping = aes(x = longitude, y = latitude, color = establishment)) + geom_point(alpha=.4) 
```

This state it holds up a lot better. It does appear that everywhere there is a Denny's, there is a Laquinta very close. Way more Laquintas though. 
